---
title: "Bonferroni Function Development"
Author: "Lynsey Keator"
Date: "April 2, 2020"
output: html_notebook
---

This notebook is for developing a function that will construct a Bonferroni confidence interval. Bonferroni takes Type 1 error rate and divides by number of comparisons


This is the bonf.ci function.
```{r}

bonf_ci <- function(the_data, the_comp, conf_level){
  
## These next 2 lines are to create vectors for both the lower and upper bounds of the 
## confidence interval.  
  
  lb <- 1:nrow(the_pairs)
  ub <- 1:nrow(the_pairs)

## Next, add these to the matric that we will return, but first copy the designation of pairs.
  
  the_cis <- cbind(the_pairs, lb, ub)

## Calculate the confidence level when we have a number of comparisons. 
## Now that I can determine the number of comparisons with the number of rows
## in the pairs matrix.
  
  pair_conf <- cbind(the_pairs, lb, ub)

## Now I loop for the number of comparisons that I am using to construct confidence 
## intervals.
  
  for (i in 1:nrow(the_pairs)){
    
## This pulls out the data that I need for the comparison on the designated row,
## which is tracked by "i" in the loop.
    
    pair <- subset(the_data, the_data[,1] == the_pairs[i, 1] |
                     the_data[,1] == the_pairs[i,2])
    
## This calculates the confidence interval.
    
    t_results <- t.test(pair[,2] ~ pair[,1], conf.level = pair_conf)
    
## These lines pull out the lower and upper bounds and add them to the matrix that I'm going to return.
    
    the_cis[i,3] <- t_results$conf.int[1]
    the_cis[i,4] <- t_results$conf.int[2]
    
  }

  return(the_cis)
}
```

This is for testing the bonf.ci function with hsb2 data. I assume that the hsb2 data are in the same directory as this notebook.

```{r}
## Outside of the function, call the here function. When done with fx development, can move function into script and keep with other functions to source in when Bonf is needed.

library(here)

hsb2 <- read.csv(here("Data", "hsb2.csv"))

## This is going to create the matrix to send to the function with my data.

my_data <- cbind(hsb2$ses, hsb2$math)

## Looking at high SES to low SES, medium SES to low SES
## Each row tells function what comparisons to make

## my_comp <- matrix(c(3, 1,
##                  2, 1),
##                  ncol = 2,
##                  byrow = TRUE)

first <- c(3,2)
second <- c(1,1)
my_pairs <- cbind(first, second)

#Obtain CI by calling function

bonf_ci(my_data, my_pairs, 0.95)

```

**Meta Coding:**
1. Function needs to provide confidence intervals (lower bound and upper bound)
2. Function will need explanatory and response variables (send in two column matrix)
  Data needed: matrix
3. Function will need to know what confidence intervals/what comparisons you want (send in two column matrix, with each row specifying the groups that need to be compared). Consider the example we have four groups (1 control and 3 treatment). These are 3 conditions (1 vs 2, 2 vs 3, 1 vs 4). In the computer's language this is equal to the number of rows.
  Data needed: matrix

1      2
1      3
1      4

*Send explanatory response variables in 1 matrix (#2), send comparisons in a second matrix (#3)*

4. nrow can tell me how many comparisons are needed
5. Function will need to know what confidence level we want
  Data needed: numeric
  
6. Function will return a matrix with the groups being compared and the lower and upper bound of the intervals
  Data needed: matrix with data for CIs + ub and lb

1, 2, lb, ub
1, 3, lb, ub
1, 4, lb, ub

Subset every row from our data that says either 3 or 1
Once data are pulled, need to do ttest with those data

*We want a matrix back from the function with the ub and lb attached to comparisons (lb added to 3rd column, ub added to 4th column)*

Of note, when we run ttest, not going to give 95%, we need to divide (and this is indicated by nrow, we need to divide by 3 to control error rate)

**Big Steps:**
Loop {

Calculate the new confidence level (alpha divided by 2, whats confidence based on that)

t.test, but need to adjust for number of comparisons were giving (nrow)

t.test(y ~ x, conf.level = 0.95)

save the upper and lower bound

}

Look at the first comparison that I want and pull out that data

Put the lower and upper bound in the matrix

Return

#### Future Enhancements
Here are some ideas for the future.
* Make is clear which way the means are subtracted.
* Be able to send a data frame instead of a matrix.
* Be able to send all the data and then indicate which variables to use in the comparisons
* Be able to use a factor instead of numbers for the explanatory variable
* Include some error checking, such as making sure the user picked groups that are actually in the data.
* Be able to do complex comparisons, not just pairwise.